Calibrating MEM seeding heuristics

### 2018-05-03 MEM.tex
About to swap v and w.

@@

By nature, reads are sequences of \texttt{A}, \texttt{T}, \texttt{G} and
\texttt{C}, but for the purpose of the seeding process, it is more useful
to consider them as a sequence of symbols telling if the nucleotides match
the target and~/ or the duplicates. Here, we will consider that reads are
sequences of the five symbols $\Downarrow$, $\downarrow$, $\uparrow$, $\sim$
and $\square$.

@@

\begin{definition}
At a given position of the read, a duplicate thread is said to be
\emph{strictly masking} if its match streak is strictly longer than the
match streak of the target. A duplicate thread is said to be
\emph{potentially masking} if its match streak is the same as the target.
\end{definition}

@@

\begin{figure}[h]
\centering
\includegraphics[scale=1]{no_3MEM_graph.pdf}
\caption{\textbf{Transfer graph of a read without 3-MEM}. 
Some legend.}
\label{fig:noMEM_graph}
\end{figure}

@@

@@

@@

\begin{remark}
When $N = 1$, $V_i = 0$ and none of the states $(\sim)_i$ is accessible.
The symbol $\sim$ can exist only if there is more than one duplicate.
\end{remark}

@@

>>>>

EXPLAIN THAT THE TAIL IS AN UNTERMINATED SEGMENT.

@@

In order to compute seeding probabilities, we need to consider that reads
are sequences of letters from the \emph{seeding alphabet} rather than from
the nucleotide alphabet. them as a sequence of symbols telling if the
nucleotides match the target and~/ or the duplicates. Here, we will
consider that reads are sequences of the five symbols $\Downarrow$,
$\downarrow$, $\uparrow$, $\sim$ and $\square$.

@@

\begin{definition}
At a given position of the read, a duplicate thread is said to be
\emph{strictly masking} if its success streak is strictly longer than the
match streak of the target. A duplicate thread is said to be
\emph{potentially masking} if its match streak is the same as the target.
\end{definition}

@@

%\begin{figure}[h]
%\centering
%\includegraphics[scale=1]{no_3MEM_graph.pdf}
%\caption{\textbf{Transfer graph of a read without 3-MEM}. 
%Some legend.}
%\label{fig:noMEM_graph}
%\end{figure}

@@

\begin{remark}
If $N = 0$, the states $(\downarrow)_2, \ldots, (\downarrow)_{\gamma-1}$
are inacessible. Consistently, the mononomial $u_i(z)$ is equal to $qz$ if
$i = 1$ and $0$ otherwise.
\end{remark}

@@

\begin{remark}
The term $\alpha_i^N-\alpha_i^N$ is the probability that the last dupicate
thread fails in one error followed by $i$ error-free nucleotides. The term
$\gamma_i^N - \delta_{i-1}^N$ is the probability that all threads fail in
$i$ error-free nucleotides...
\end{remark}

@@

\begin{remark}
When $N = 1$, $V_i = 0$ and the states $(\sim)_i$ are inaccessible.
The symbol $\sim$ can exist only if there is more than one duplicate.
\end{remark}

@@

In the case $N = 1$, reads never contain the symbol $\sim$ and the matrix
is
\begin{equation*}
\begin{blockarray}{cccccc}
   & \DN & \dn & \up{1} & \ldots & \up{\gamma-1} \\
\begin{block}{c[ccccc]}
\DN & A_\gamma(z) & \tilde{A}_\infty(z) & u_1(z)
    & \ldots & u_{\gamma-1}(z) \\
\dn & B_\infty(z) & \tilde{B}_\infty(z) & w_1(z)
    & \ldots & w_{\gamma-1}(z) \\
\up{1} & D_{\gamma-1}(z) & \tilde{D}_{\gamma-1}(z) & 0 &
    \ldots & 0 \\
\up{2} & D_{\gamma-2}(z) & \tilde{D}_{\gamma-2}(z) & 0 &
    \ldots & 0  \\
\vdots & \vdots & \vdots & \vdots & \ddots & \vdots \\
\up{\gamma-1} & D_1(z) & \tilde{D}_1(z) & 0 & \ldots & 0 \\
\end{block}
\end{blockarray}
\end{equation*}

In the case $N = 0$, the symbol $\downarrow$ also disappears. Some
polynomials have a particularly simple form in this case, and the
body of the transfer matrix is
\begin{equation*}
\begin{blockarray}{ccc}
   & \DN & \up{1} \\
\begin{block}{c[cc]}
\DN    & pz              & qz \\
\up{1} & D_{\gamma-1}(z) & 0  \\
\end{block}
\end{blockarray}
\end{equation*}
